Advancing Renewable Gas Storage using Flat-histogram Methods

### 2020-06-06 dissertation/introduction.tex
finished all of the corrections and edits.

Monte Carlo methods are a powerful simulation tool that allow scientists to
explore properties of systems that may otherwise by inaccessible via
experimental methods. Monte Carlo methods can be used to solve a variety of
physical problems yielding insight into statistical averages due to random
processes. During the last two decades considerable research has been done on
developing and refining flat-histogram Monte Carlo methods. These Monte Carlo
(MC) methods are capable of exploring all of the energy phase space for all
given temperatures in a \emph{single} simulation. In this work, we first
outline the necessary statistical mechanics background. We provide a
description of MC methods and flat-histogram methods, as well as, our newly
developed method SAD (stochastic approximation Monte Carlo with a dynamic
update factor). We discuss physics rich test systems that can be used to
explore the convergence and statistical properties of our newly developed
method SAD.

A driving motivation for developing novel Monte Carlo methods that have physically based tunable parameters is to provide simulation methods for exploring metal-organic frameworks (MOFs). In this work, we examine MOFs and
their application to gas storage and delivery. We formulate a theoretical upper
bound on the maximum amount of gas that can be stored and released at different
pressures. We propose SAD as a powerful method for examining thermodynamic properties for MOFs.  We lay a groundwork for 2D SAD which would be able to compute isotherms for any given MOF thereby providing a powerful way for researchers to determine a MOF's suitability for gas storage applications.

In the following chapters, we follow the manuscript outline with each chapter corresponding to our published or submitted research on the SAD method applied to various physical systems and our work on establishing a theoretical upper bound on gas storage and delivery for MOFs.

@@

\subsection{Definition of Entropy $S(E)$}\label{entropy}
The entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states using one of the most important formulae in physics:
\begin{align}\label{sdos}
    S(E,V,N) \equiv k_B \ln\Omega(E,V,N)
\end{align}
where the entropy is defined as the logarithm of the density of states and $k_B$ is the Boltzmann constant. An important aspect of this definition in terms of logarithms is that the entropy for multiple isolated systems is additive.
\begin{align}
    S(E_1,E_2) &= k_B \ln(\Omega(E_1)\Omega_2(E_2)) \\
    &= k_B \left( \ln(\Omega_1(E_1)) + \ln(\Omega_2(E_2))\right) \\
    &= k_B \left( S_1(E_1) + S_2(E_2)\right)
\end{align}
From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by referring to Eq.~\HR{sdos}.
\begin{align}
    \frac{1}{T} = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
\end{align}
The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in Section~\HR{laws}.

\subsection{The Laws of Thermodynamics and $S(E)$}\label{laws}
The \nth{1} law of thermodynamics states that for an isolated system $\Gamma$, the total internal energy is equal to the sum of the heat $Q$ and work $W$ done.
\begin{align}
    \D{U} = \D{Q} + \D{W}
\end{align}
The \nth{1} law follows directly from conservation of energy. The thermodynamic identity states that for any infinitesimal change in the system and fixed number of particles $N$, the internal energy can be written in terms of the entropy.
\begin{align}
    \D{U} =  T\D{S} - p\D{V} \\
    T = \left(\pdv{U}{S}\right)_V
\end{align}

The \nth{2} law of thermodynamics states that for any isolated system $\Gamma$ the entropy $S(E) \geq 0$. The heat of the system can also be related to the entropy using the relation:
\begin{align}
    \delta Q = T\delta S
\end{align}
The relationship between heat and entropy for a system is invaluable for the \nth{1} law to relate the conjugate variables T, $S(E)$, and $U(E)$.

@@

The probability $\mathcal{P}$ of selecting a given micro-state at random is
\begin{align}
    \mathcal{P} = e^{\nicefrac{\left(F-E\right)}{k_BT}}
\end{align}
where $F\left(N,V,T\right)$ is the free energy of the system $\Gamma$ and will be discussed in detail in Section~\HR{potentials}.

\subsection{Boltzmann Factor and Partition Function}
In statistical thermodynamics $\beta$ is defined as the coldness or the reciprocal of the temperature.
\begin{align}
    \beta = \frac{1}{k_BT} = \frac{1}{k_B}
\end{align}
where $k_B$ is the Boltzmann constant. The partition function for a canonical ensemble can then be written in terms of the Boltzmann factor where $E$ is the total energy of the system in the respective microstate at index $i$.
\begin{align}
    Z = \sum_i e^{-\beta E_i}
\end{align}

\subsection{Thermodynamic Potentials}\label{potentials}
Thermodynamic potentials describe a given thermodynamic state of a given system and
have units of energy $k_BT$. In this section, we will discuss five common thermodynamic potentials and how they are mathematically related through Maxwell relations.
While the internal energy $U(E)$ has been briefly mentioned in Section~\HR{laws}, we outline in detail the relationship between the state variables.
\begin{align}
    \D{U(S,V,N)} = T\D{S}-p\D{V}+\mu\D{N}
\end{align}
We can write the temperature $T$, pressure $p$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxU}
    T = \left( \pdv{U}{S}\right )_{V,N}  \quad\quad 
    p = -\left( \pdv{U}{V}\right )_{S,N}  \quad\quad 
    \mu = \left( \pdv{U}{N}\right )_{S,V}  \quad
\end{align}
The Helmholtz free energy $F(T,V,N)$ can be written as follows:
\begin{align}
    \D{F(T,V,N)} = \D{(U-TS)} = -S\D{T}-p\D{V}+\mu\D{N}
\end{align}
We can write the entropy $S$, pressure $p$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxF}
    S = -\left( \pdv{F}{T}\right )_{V,N}  \quad\quad 
    p = \left( \pdv{F}{V}\right )_{T,N}  \quad\quad 
    \mu = \left( \pdv{F}{N}\right )_{T,V}  \quad
\end{align}
The Enthalpy $H(S,p,N)$ can be written as follows:
\begin{align}
    \D{H(S,p,N)} = \D{(U+pV)} = -T\D{S}+V\D{p}+\mu\D{N}
\end{align}
We can write the temperature $T$, volume $V$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxH}
    T = \left( \pdv{H}{S}\right )_{p,N}  \quad\quad 
    V = -\left( \pdv{H}{p}\right )_{S,N}  \quad\quad 
    \mu = \left( \pdv{H}{N}\right )_{S,p}  \quad
\end{align}
The Gibbs free energy $G(T,p,N)$ can be written as follows:
\begin{align}
    \D{G(T,p,N)} = \D{(U+pV-TS)} = -S\D{T}+V\D{p}+\mu\D{N}
\end{align}
We can write the entropy $S$, volume $V$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxG}
    S = \left( \pdv{G}{T}\right )_{p,N}  \quad\quad 
    V = -\left( \pdv{G}{p}\right )_{T,N}  \quad\quad 
    \mu = \left( \pdv{G}{N}\right )_{T,p}  \quad
\end{align}

\subsection{Maxwell Relations}\label{maxwell}
Maxwell's relations are second derivative equalities that relate important thermodynamic variables. Maxwell relations are an extension of the Schwarz theorem which states that the order of differentiation is irrelevant.
\begin{align}
    \pdv{}{\theta}\left(\pdv{\Theta}{\phi}\right)=\pdv{}{\phi}\left(\pdv{\Theta}{\theta}\right)
\end{align}
From this theorem and Section~\HR{potentials}, we can derive the most common of Maxwell's relations. The relation involving internal energy $U\left(S,V,N\right)$ relates the natural variables of pressure $p$ and temperature $T$ and can be derived using Equation~\ref{maxU}.
\begin{align}
    \left( {\pdv{U}{S}{V}} \right )_{N} = \left( \pdv{T}{V}\right )_{S,N} = - \left( \pdv{p}{S}\right )_{V,N}
\end{align}
The relation involving the Helmholtz free energy $F\left(T,V,N\right)$ relates the natural variables of entropy $S$ and pressure $p$ and can be derived using Equation~\ref{maxF}.
\begin{align}
    \left( {\pdv{F}{T}{V}} \right )_{N} = \left( \pdv{S}{V}\right )_{T,N} = \left( \pdv{p}{T}\right )_{V,N}
\end{align}
The relation involving the Enthalpy $H\left(S,p,N\right)$ relates the natural variables of temperature $T$ and volume $V$ and can be derived using Equation~\ref{maxH}.
\begin{align}
    \left( {\pdv{H}{S}{p}} \right )_{N} = \left( \pdv{T}{p}\right )_{S,N} =  \left( \pdv{V}{S}\right )_{p,N}
\end{align}
The relation involving the Gibbs free energy $G\left(T,p,N\right)$ relates the natural variables of entropy $S$ and volume $V$ and can be derived using Equation~\ref{maxG}.
\begin{align}
    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = \left( \pdv{V}{T}\right )_{p,N}
\end{align}

\subsection{Thermodynamic Properties}\label{thermoprop}
The Maxwell relations from Section~\HR{maxwell} can be written in terms of thermodynamic properties of the material.  The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature by a given amount.
\begin{align}
    C_V = T\left(\pdv{S}{T}\right )_{V,N}
\end{align}
The isobaric heat capacity $C_p$, which is always greater than or equal to $C_V$ according to the Mayer relation, can also be written in terms of the entropy $S$ and temperature $T$ of the system.
\begin{align}
    C_p = T\left(\pdv{S}{T}\right )_{p,N}
\end{align}
The Mayer relation gives the difference of the isobaric and isochoric heat capacities in terms of the thermal expansion coefficient and the isothermal compressibility.
\begin{align}
    C_p - C_V = VT\frac{\alpha^2}{\kappa_T}
\end{align}
The isobaric coefficient of thermal expansion determines to what extent a given material will deform as a result of a change in temperature.
\begin{align}
    \alpha_p = \frac{1}{V}\left( \pdv{V}{T} \right )_{p,N}
\end{align}
Similarly, the isochoric coefficient of thermal expansion can be written as the following:
\begin{align}
    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}
\end{align}
The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written in terms of the following:
\begin{align}
    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}
\end{align}
Similarly, the isentropic compressibility can be written as the following:
\begin{align}
    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}
\end{align}
A major driving force behind introducing the concepts of thermal expansion and compressibility is to right complex derivatives in terms of these parameters. The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{U}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} - \alpha TV \quad\quad
    \left( \pdv{S}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha T} + \alpha V
\end{align}
Additionally, the enthalpy derivatives can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{H}{T} \right )_{V,N} = C_p + \frac{\alpha V}{\kappa_T}\left(1-\alpha T\right) \quad\quad
    \left( \pdv{H}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} + V\left(1-\alpha T\right)
\end{align}
Finally, the Gibbs free energy derivatives can be written in terms of the isothermal compressibility, the thermal coefficient of expansion, and the entropy.
\begin{align}
    \left( \pdv{G}{T} \right )_{V,N} = \frac{\alpha V}{\kappa_T}-S \quad\quad
    \left( \pdv{G}{p} \right )_{V,N} = V-\frac{\kappa_T S}{\alpha}
\end{align}

\section{Grand Canonical Ensemble}
The grand canonical ensemble is a representation for a system $\Gamma$ that is determined by a fixed chemical potential $\mu$, a fixed volume $V$, and a fixed temperature $T$. All possible states for the system $\Gamma\left(\mu,V,T\right)$ have the same chemical potential, volume, and temperature due to the system being in thermal and chemical equilibrium with a given resevoir.
The grand canonical ensemble is for this reason also known as the $\mu VT$ ensemble.

@@

\subsection{Chemical Potential $\mu$}
The chemical potential $\mu$ of a given system $\Gamma$ is representative of the energy that can be gained or lost due (often because of a phase transition) to a change in the conjugate variable particle number $N$. Analogous to potential energy, higher chemical potential drives particles to lower potential (thereby usually changing the particle number). The chemical potential $\mu_i$ of species $i$ can be represented using a variety of free energies (all of which are equivalent but not necessarily useful for any given experiment).
\begin{align}
    \mu_i = \left( \pdv{G}{N_i}\right )_{T,P,N_{j \neq i}}  \quad
    \mu_i = \left( \pdv{H}{N_i}\right )_{S,P,N_{j \neq i}}  \quad
    \mu_i = \left( \pdv{F}{N_i}\right )_{T,V,N_{j \neq i}}  
\end{align}

@@

\section{Monte Carlo Simulation}\label{monte}
Experiment and simulation blend together to help form a more complete understanding of our knowledge of how the universe works. Computational `experiments' such as Monte Carlo methods aid in this understanding by allowing scientists to explore system configurations that are not easily accessible to traditional experiment.

@@

The motivation for new research into developing MC methods stems from the desire to solve ever-increasingly complex physical problems in a faster simulation time. Many of the originally developed MC methods are unable to effectively explore the complex phase space of real (and even many test) systems. Flat-histogram methods (introduced in Section~\HR{fhm}) are able to explore all energies and temperatures in a `single' simulation. In addition, they are able to explore complex energy landscapes without getting `stuck' (either never converging or converging at such a slow rate as to not practically converge). Many of these flat-histogram methods also have proven convergence on the statistical error thus guaranteeing convergence.
In the following sections, we will discuss a number of MC methods including importance sampling canonical methods, and flat-histogram methods leading up to our method SAD introduced in this work discussed in Section~\HR{fhm}.

@@

There are a variety of PRNGs each with benefits and few without some form of drawback. We detail some of the well known PRNGs here. Linear congruential generators (LCGs) have a very fast time performance and are straightforward to program (in terms of size and complexity). Unfortunately, LCGs have multiple issues regarding the generated random number statistical quality~\cite{coddington1994analysis, barry1996recommendations}. While the period of the PRNG can be made acceptable by using 64-bit integers, the statistical correlations remain. Other PRNGs are the belong to the shift-register family (Xorshift RNGs). Like the LCG family, the shift-register RNGs fail many statistical tests~\cite{panneton2005xorshift} and are therefore often considered unacceptable for use in Monte Carlo simulations.

In our work, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate). The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and incredible statistical quality~\cite{blackman2018scrambled}.

\subsection{Canonical Monte Carlo}\label{cmc}
In this section, we introduce canonical Monte Carlo and how it relates to flat-histogram methods introduced in Section~\HR{fhm}. The first importance sampling method developed was the famous Metropolis method~\cite{metropolis1953equation}.  The method works by choosing an initial state for the system $\Gamma$. Next, a random site $\Gamma_i$ is chosen and a move is proposed that will result in an energy change $\Delta E$ if accepted. The acceptance of the move is determined by whether a random number chosen on a range from zero to one is less than a Boltzmann factor. The process of moves and random number generation is repeated until some termination condition is met. This canonical method performs calculations at a fixed temperature $T$ only. Re-weighting can extrapolate in the vicinity of the simulated temperature, but this approach is limited. Another limitation of canonical Monte Carlo is that for complex energy landscapes and/or large systems the simulation can undergo a phenomena known as `critical slowing down' as the simulation will never or almost never visit the ground state energy $E_{\min}$~\cite{berg2000introduction}. These limitations are what drive the need for methods that can avoid getting `stuck' and that can simulate all temperatures in a single simulation.

\subsection{Flat-Histogram Methods}\label{fhm}
The goal of flat-histogram methods (also called \emph{broad-histogram}
or \emph{multicanonical} methods) is to simulate each energy with
similar accuracy.  In this way, they can accurately determine the density of states
over a broad range of energies and thus determine the thermodynamic
quantities such as heat capacity or internal energy
over a broad range of temperatures.
Properties that require more information, such as a spatial
correlation function or a response function, can still be computed
for any temperature, provided statistics are collected for each
individual energy, which can then be reweighted for any
temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
errington2003direct}.
All the flat-histogram Monte Carlo methods begin with randomly chosen
`moves' which change the state of the system and must satisfy
detailed balance.  All of the tested algorithms differ in how they determine the
probability of accepting a move and in what additional statistics must
be collected in order to decide on that probability.
Flat-histogram methods calculate the density of states $D(E)$ for a discrete set
of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
trebst2004optimizing}. Therefore, energy binning becomes an important
consideration for systems with a continuum of possible energies.  Energy bins
are typically of uniform size for the entire energy
continuum~\cite{fasnacht2004adaptive}.

>>>>

Monte Carlo (MC) methods are a powerful simulation tool that allow scientists to
explore properties of systems that may otherwise by inaccessible via
experimental methods. Monte Carlo methods can be used to solve a variety of
physical problems yielding insight into statistical averages due to random
processes. During the last two decades considerable research has been done on
developing and refining flat-histogram Monte Carlo methods. These Monte Carlo
methods are capable of exploring all of the energy phase space for all
given temperatures in a \emph{single} simulation. In this work, we first
outline the necessary statistical mechanics background. We provide a
description of MC methods and flat-histogram methods, as well as, our newly
developed method SAD (stochastic approximation Monte Carlo with a dynamic
update factor). We discuss physics rich test systems that can be used to
explore the convergence and statistical properties of our newly developed
method SAD.

A driving motivation for developing novel Monte Carlo methods that have
physically based tunable parameters is to provide simulation methods for
exploring porous materials. In this work, we examine the theoretical upper
bound on the storage capacity of porous materials in the context of
light-vehicle fuel storage. We formulate a theoretical upper bound on the
maximum amount of gas that can be stored and released at different pressures.
We propose SAD as a potentially powerful method for examining thermodynamic
properties for porous materials. This body of research serves as a groundwork
for multidimensional SAD which would be able to compute isotherms for any given
porous material thereby providing a powerful way for researchers to determine
the suitability for gas storage applications.

In the following chapters, we follow the manuscript outline with each chapter
corresponding to our published or submitted research on the SAD method applied
to various physical systems, and our work on establishing a theoretical upper
bound on gas storage and delivery for MOFs.

@@

\subsection{The Boltzmann Entropy $S(E)$}\label{entropy}
The Boltzmann entropy $S(E)$ can be defined for any given system $\Gamma$ with discrete energy states by relating it to the microstates:
\begin{align}\label{sdos}
    S(E,V,N) \equiv k_B \ln\Omega(E,V,N)
\end{align}
where the entropy is defined as the logarithm of the density of states and $k_B$ is the Boltzmann constant. An important aspect of this definition in terms of logarithms is that the entropy for multiple isolated systems is additive.
\begin{align}
    S(E_1,E_2) &= k_B \ln(\Omega(E_1)\Omega_2(E_2)) \\
    &= k_B \left( \ln(\Omega_1(E_1)) + \ln(\Omega_2(E_2))\right) \\
    &= \left( S_1(E_1) + S_2(E_2)\right)
\end{align}
From the \nth{1} law of thermodynamics, the system temperature can be written in terms of the density of states by referring to Eq.~\HR{sdos}.
\begin{align}
    \frac{1}{T} = \left(\diff{S(E)}{E}\right)_V = \frac{k_B}{\Omega(E)}\left(\diff{\Omega(E)}{E}\right)_V
\end{align}
The relationship between entropy $S(E)$, $\Omega(E)$, and $T$ will be discussed in further depth in Section~\HR{laws}.

\subsection{Thermodynamics in the Micro-Canonical Ensemble}\label{laws}
The statistical entropy defined in Eq.~\HR{sdos} is consistent with the \nth{1} and \nth{2} laws of
thermodynamics for any system with uniformly distributed microstates.
The \nth{1} law of thermodynamics states that for an isolated system $\Gamma$, the total internal energy is equal to the sum of the heat $Q$ and work $W$ done.
\begin{align}
    \D{U} = \D{Q} + \D{W}
\end{align}
The \nth{1} law follows directly from conservation of energy. The thermodynamic identity states that for any infinitesimal change in the system and fixed number of particles $N$, the internal energy can be written in terms of the entropy.
\begin{align}
    \D{U} =  T\D{S} - p\D{V} \\
    T = \left(\pdv{U}{S}\right)_V
\end{align}

The \nth{2} law of thermodynamics states that for any isolated system $\Gamma$ the entropy $S(E) \geq 0$. The heat of the system can also be related to the entropy using the relation:
\begin{align}
    \delta Q = T\delta S
\end{align}
The relationship between heat and entropy for a system is useful for the \nth{1} law to relate the conjugate variables T, $S(E)$, and $U(E)$.

@@

The probability $\mathcal{P}$ of selecting a given micro-state $i$ at random is
\begin{align}
    \mathcal{P_i} = \frac{e^{\nicefrac{\left(-E_i\right)}{k_BT}}}{Z}
\end{align}
where $Z$ is the canonical partition function and will be discussed in detail in Section~\HR{betaZ}.

\subsection{The Canonical Partition Function}\label{betaZ}
In statistical thermodynamics $\beta$ is defined as the coldness or the reciprocal of the canonical temperature.
\begin{align}
    \beta = \frac{1}{k_BT}
\end{align}
where $k_B$ is the Boltzmann constant. The partition function for a canonical ensemble can then be written in terms of the Boltzmann factor where $E$ is the total energy of the system in the respective microstate at index $i$.
\begin{align}
    Z = \sum_i e^{-\beta E_i}
\end{align}
The partition function depends both on the microstate energies as well as the
temperature $T$. Because the microstate energies are influenced by microscopic
properties, $Z$ has important statistical meaning.

\section{Thermodynamic Potentials}\label{potentials}
Thermodynamic potentials describe a given thermodynamic state of a given system
and have units of energy $k_BT$. In this section, we will discuss the
thermodynamic potentials ($U$, $F$, $H$, and $G$) and how they are mathematically related through
Maxwell relations. While the internal energy $U(E)$ has been briefly mentioned
in Section~\HR{laws}, we outline in detail the relationship between the state
variables.
\begin{align}
    \D{U(S,V,N)} = T\D{S}-p\D{V}+\mu\D{N}
\end{align}
We can write the temperature $T$, pressure $p$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxU}
    T = \left( \pdv{U}{S}\right )_{V,N}  \quad\quad 
    p = -\left( \pdv{U}{V}\right )_{S,N}  \quad\quad 
    \mu = \left( \pdv{U}{N}\right )_{S,V}  \quad
\end{align}
The Helmholtz free energy $F(T,V,N)$ can be written as follows:
\begin{align}
    \D{F(T,V,N)} = \D{(U-TS)} = -S\D{T}-p\D{V}+\mu\D{N}
\end{align}
We can write the entropy $S$, pressure $p$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxF}
    S = -\left( \pdv{F}{T}\right )_{V,N}  \quad\quad 
    p = \left( \pdv{F}{V}\right )_{T,N}  \quad\quad 
    \mu = \left( \pdv{F}{N}\right )_{T,V}  \quad
\end{align}
The Enthalpy $H(S,p,N)$ can be written as follows:
\begin{align}
    \D{H(S,p,N)} = \D{(U+pV)} = -T\D{S}+V\D{p}+\mu\D{N}
\end{align}
We can write the temperature $T$, volume $V$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxH}
    T = \left( \pdv{H}{S}\right )_{p,N}  \quad\quad 
    V = -\left( \pdv{H}{p}\right )_{S,N}  \quad\quad 
    \mu = \left( \pdv{H}{N}\right )_{S,p}  \quad
\end{align}
The Gibbs free energy $G(T,p,N)$ can be written as follows:
\begin{align}
    \D{G(T,p,N)} = \D{(U+pV-TS)} = -S\D{T}+V\D{p}+\mu\D{N}
\end{align}
We can write the entropy $S$, volume $V$, and chemical potential $\mu$ in terms of derivatives of state variables.
\begin{align}\label{maxG}
    S = \left( \pdv{G}{T}\right )_{p,N}  \quad\quad 
    V = -\left( \pdv{G}{p}\right )_{T,N}  \quad\quad 
    \mu = \left( \pdv{G}{N}\right )_{T,p}  \quad
\end{align}

\subsection{Maxwell Relations}\label{maxwell}
Maxwell's relations are second derivative equalities that relate thermodynamic variables. Maxwell relations are formalism of the Schwarz theorem which states that the order of differentiation is irrelevant.
\begin{align}
    \pdv{}{\theta}\left(\pdv{\Theta}{\phi}\right)=\pdv{}{\phi}\left(\pdv{\Theta}{\theta}\right)
\end{align}
From this theorem and Section~\HR{potentials}, we can derive the most common of Maxwell's relations. The relation involving internal energy $U\left(S,V,N\right)$ relates the natural variables of pressure $p$ and temperature $T$ and can be derived using Equation~\ref{maxU}.
\begin{align}
    \left( {\pdv{U}{S}{V}} \right )_{N} = \left( \pdv{T}{V}\right )_{S,N} = - \left( \pdv{p}{S}\right )_{V,N}
\end{align}
The relation involving the Helmholtz free energy $F\left(T,V,N\right)$ relates the natural variables of entropy $S$ and pressure $p$ and can be derived using Equation~\ref{maxF}.
\begin{align}
    \left( {\pdv{F}{T}{V}} \right )_{N} = \left( \pdv{S}{V}\right )_{T,N} = \left( \pdv{p}{T}\right )_{V,N}
\end{align}
The relation involving the Enthalpy $H\left(S,p,N\right)$ relates the natural variables of temperature $T$ and volume $V$ and can be derived using Equation~\ref{maxH}.
\begin{align}
    \left( {\pdv{H}{S}{p}} \right )_{N} = \left( \pdv{T}{p}\right )_{S,N} =  \left( \pdv{V}{S}\right )_{p,N}
\end{align}
The relation involving the Gibbs free energy $G\left(T,p,N\right)$ relates the natural variables of entropy $S$ and volume $V$ and can be derived using Equation~\ref{maxG}.
\begin{align}
    \left( {\pdv{G}{T}{p}} \right )_{N} = -\left( \pdv{S}{p}\right )_{T,N} = \left( \pdv{V}{T}\right )_{p,N}
\end{align}

\subsection{Thermodynamic Properties}\label{thermoprop}
The Maxwell relations from Section~\HR{maxwell} can be written in terms of thermodynamic properties of the material.  The isochoric heat capacity $C_V$ of a given material is defined as the amount of heat necessary to raise the temperature of the material.
\begin{align}
    C_V = T\left(\pdv{S}{T}\right )_{V,N}
\end{align}
The isobaric heat capacity $C_p$, which is always greater than or equal to $C_V$ according to the Mayer relation, can also be written in terms of the entropy $S$ and temperature $T$ of the system.
\begin{align}
    C_p = T\left(\pdv{S}{T}\right )_{p,N}
\end{align}
The Mayer relation gives the difference of the isobaric and isochoric heat capacities in terms of the thermal expansion coefficient and the isothermal compressibility.
\begin{align}
    C_p - C_V = VT\frac{\alpha^2}{\kappa_T}
\end{align}
The isobaric coefficient of thermal expansion determines to what extent a given material will deform as a result of a change in temperature.
\begin{align}
    \alpha_p = \frac{1}{V}\left( \pdv{V}{T} \right )_{p,N}
\end{align}
Similarly, the isochoric coefficient of thermal expansion can be written as
\begin{align}
    \alpha_V = \frac{1}{p}\left( \pdv{p}{T} \right )_{V,N}.
\end{align}
The compressibility determines to what extent a given material will change volume based on the pressure applied to the system.  The isothermal compressibility can be written:
\begin{align}
    \kappa_T = -\frac{1}{V}\left( \pdv{V}{p} \right )_{T,N}.
\end{align}
Similarly, the isentropic compressibility can be written as
\begin{align}
    \kappa_S = -\frac{1}{V}\left( \pdv{V}{p} \right )_{S,N}.
\end{align}
A major driving force behind introducing the concepts of thermal expansion and compressibility is the ability to write complex derivatives in terms of these parameters. The internal energy derivative and entropy derivative can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{U}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} - \alpha TV \quad\quad
    \left( \pdv{S}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha T} + \alpha V
\end{align}
Additionally, the enthalpy derivatives can be written in terms of both the isothermal compressibility and the thermal coefficient of expansion.
\begin{align}
    \left( \pdv{H}{T} \right )_{V,N} = C_p + \frac{\alpha V}{\kappa_T}\left(1-\alpha T\right) \quad\quad
    \left( \pdv{H}{p} \right )_{V,N} = \frac{\kappa_T C_p}{\alpha} + V\left(1-\alpha T\right)
\end{align}
Finally, the Gibbs free energy derivatives can be written in terms of the isothermal compressibility, the thermal coefficient of expansion, and the entropy.
\begin{align}
    \left( \pdv{G}{T} \right )_{V,N} = \frac{\alpha V}{\kappa_T}-S \quad\quad
    \left( \pdv{G}{p} \right )_{V,N} = V-\frac{\kappa_T S}{\alpha}
\end{align}

\section{Grand Canonical Ensemble}
The grand canonical ensemble is a representation for a system $\Gamma$ that is determined by a fixed chemical potential $\mu$, a fixed volume $V$, and a fixed temperature $T$. All possible states for the system $\Gamma\left(\mu,V,T\right)$ have the same chemical potential, volume, and temperature due to the system being in thermal and chemical equilibrium with a given reservoir.
The grand canonical ensemble is for this reason also known as the $\mu VT$ ensemble.

@@

\subsection{Chemical Potential $\mu$}
The chemical potential $\mu$ of a given system $\Gamma$ is representative of the energy that can be gained or lost due (often because of a phase transition) to a change in the conjugate variable particle number $N$. Analogous to potential energy, higher chemical potential drives particles to lower potential (thereby usually changing the particle number). The chemical potential $\mu_i$ of species $i$ can be represented using a variety of free energies (all of which are equivalent).
\begin{align}
    \mu_i = \left( \pdv{G}{N_i}\right )_{T,P,N_{j \neq i}}  \quad
    \mu_i = \left( \pdv{H}{N_i}\right )_{S,P,N_{j \neq i}}  \quad
    \mu_i = \left( \pdv{F}{N_i}\right )_{T,V,N_{j \neq i}}  
\end{align}

@@

\section{Monte Carlo Simulation}\label{monte}
Experiment and simulation are both necessary to help form a more complete
understanding of our knowledge of how the universe works. Computational
`experiments' such as Monte Carlo methods aid in this understanding by allowing
scientists to explore system configurations that are not easily accessible to
traditional experiment.

@@

The motivation for new research into developing MC methods stems from the
desire to solve ever-increasingly complex physical problems in a faster
simulation time. Many of the originally developed MC methods are unable to
effectively explore the complex phase space of real (and even many test)
systems. Flat-histogram methods (introduced in Section~\HR{fhm}) are able to
explore all energies and temperatures in a `single' simulation. In addition,
they are able to explore complex energy landscapes without getting `stuck'
(either never converging or converging at such a slow rate as to not
practically converge). Many of these flat-histogram methods also have proven
convergence on the statistical error thus guaranteeing convergence. In the
following sections, we will discuss a number of MC methods including importance
sampling and flat-histogram methods introduced in Section~\HR{cmc} and
~\HR{fhm}.

@@

There are a variety of PRNGs each with benefits and few without some form of
drawback. We detail some of the well known PRNGs here. Linear congruential
generators (LCGs) have a very fast time performance and are straightforward to
program (in terms of size and complexity). Unfortunately, LCGs have multiple
issues regarding the generated random number statistical
quality~\cite{coddington1994analysis, barry1996recommendations}. While the
period of the PRNG can be made acceptable by using 64-bit integers, the
statistical correlations remain. Other PRNGs belong to the
shift-register family (Xorshift RNGs). Like the LCG family, the shift-register
RNGs fail many statistical tests~\cite{panneton2005xorshift} and are therefore
often considered unacceptable for use in Monte Carlo simulations.

In our work, we use the \textbf{Xoroshiro128+} RNG which is named after its logical operations (XOR $\xrightarrow{}$ rotate $\xrightarrow{}$ shift $\xrightarrow{}$ rotate). The RNG has ultra-fast speed 0.72 $\nicefrac{\text{ns}}{\text{64-bits}}$ and excellent statistical quality~\cite{blackman2018scrambled}.

\subsection{Canonical Monte Carlo}\label{cmc}
In this section, we introduce canonical Monte Carlo and how it relates to flat-histogram methods introduced in Section~\HR{fhm}. The first importance sampling method developed was the Metropolis method~\cite{metropolis1953equation}.  The method works by choosing an initial state for the system $\Gamma$. Next, a random site $\Gamma_i$ is chosen and a move is proposed that will result in an energy change $\Delta E$ if accepted. The acceptance of the move is determined by whether a random number chosen on a range from zero to one is less than a Boltzmann factor. The process of moves and random number generation is repeated until some termination condition is met. This canonical method performs calculations at a fixed temperature $T$ only. Re-weighting can extrapolate in the vicinity of the simulated temperature, but this approach is limited. Another limitation of canonical Monte Carlo is that for complex energy landscapes and/or large systems the simulation can undergo a phenomena known as `critical slowing down' as the simulation will never or almost never visit the ground state energy $E_{\min}$~\cite{berg2000introduction}. These limitations are what drive the need for methods that can avoid getting `stuck' and that can simulate all temperatures in a single simulation.
